---
title: "Building database"
author: "Priscila Stisman"
format: html
editor: visual
---

### Building Dataset

This code uses the `nyt-metadata.csv` file, which contains over 2.1 million New York Times headlines. I extracted samples from the years 2021 to 2025 to analyze how the narrative around energy policy evolved during this period—particularly before and after the Russia-Ukraine war, which began in February 2022.

The code below outlines the process used to sample headlines from each of these years and combine them into a single dataset, saved as `nyt_combined_2021_2025.csv` in the data folder.

```{r}
###Setup
library(tidyverse)
library(dplyr)
library(data.table)
library(here)
library(stringr)
```

```{r}

### Building the database from scratch 

set.seed(2710)

# Sample 2021
nyt_2021 <- fread(here("data", "nyt-metadata.csv"),
                 select = c("pub_date", "headline", "abstract", "keywords", "lead_paragraph")) %>%
  # Use ymd_hms to parse dates in YYYY-MM-DD HH:MM:SS format
  mutate(date_parsed = lubridate::ymd_hms(pub_date)) %>%
  # Filter by year
  filter(year(date_parsed) == 2021)

write_csv(nyt_2021, here("data","nyt_2021.csv"))


# Sample 2022
nyt_2022 <- fread(here("data", "nyt-metadata.csv"),
                 select = c("pub_date", "headline", "abstract", "keywords", "lead_paragraph")) %>%
  # Use ymd_hms to parse dates in YYYY-MM-DD HH:MM:SS format
  mutate(date_parsed = lubridate::ymd_hms(pub_date)) %>%
  # Filter by year
  filter(year(date_parsed) == 2022)

write_csv(nyt_2022, here("data","nyt_2022.csv"))


# Sample 2023
nyt_2023 <- fread(here("data", "nyt-metadata.csv"),
                 select = c("pub_date", "headline", "abstract", "keywords", "lead_paragraph")) %>%
  # Use ymd_hms to parse dates in YYYY-MM-DD HH:MM:SS format
  mutate(date_parsed = lubridate::ymd_hms(pub_date)) %>%
  # Filter by year
  filter(year(date_parsed) == 2023) 

write_csv(nyt_2023, here("data","nyt_2023.csv"))


# Combine them all into one dataframe
nyt_combined <- bind_rows(nyt_2021, nyt_2022, nyt_2023)
write_csv(nyt_combined, here("data", "nyt_combined_2021_2023.csv"))

```

```{r}
### Reading the csv and combining them all into a single one

# First read all your individual CSV files
nyt_2021 <- read_csv(here("data", "nyt_2021.csv"))
nyt_2022 <- read_csv(here("data", "nyt_2022.csv"))
nyt_2023 <- read_csv(here("data", "nyt_2023.csv"))


# Combine them all into one dataframe
nyt_combined_2021_2023 <- bind_rows(nyt_2021, nyt_2022, nyt_2023)

# Write the combined dataset to a new file
write_csv(nyt_combined, here("data", "nyt_combined_2021_2023.csv"))

```
